### 2016-09-19 eg6-letters/analysis-note-v2/data_analysis.tex
merge files

\chapter{Particles reconstruction and simulation}
~~~~The final state of a coherent (incoherent) DVCS event consists of three 
particles: an electron, a $^4$He (a proton), and a real photon. To identify the 
DVCS events, we first identify, individually, the different particles of 
interest. Then, events with three detected final-state particles will be 
further filtered by imposing the energy-momentum conservation laws, as will be 
presented in the following chapter.\\

Even after imposing the conservation laws, the DVCS sample will not have 
100$\%$ truly DVCS events. In our kinematical region, the main contamination to 
our DVCS channels comes from the electroproduction of neutral pions. For 
instance, in the coherent $\pi^0$-electroproduction, when one of the two-photon 
decay of the $\pi^{0}$ passes the DVCS requirements, it will be counted as a 
DVCS event. Thus, these events have to be subtracted before looking to any DVCS 
observable. For this purpose, we perform a technique in which we combine the 
measured exclusive $\pi^0$-electroproduction data sample with a Monte-Carlo 
simulation to evaluate the background in the selected DVCS sample. For this 
technique, we need to identify the experimental $\pi^0$s that come in the 
coherent and the incoherent $\pi^0$-electroproduction channels.\\

@@

\subsection{Electron identification} \label{Electron_identification}
~~~~The electron detection triggers the data acquisition system to record data 
from all the sub-detectors of CLAS. In this analysis, a particle which passes 
the following set of criteria is assumed to be a good electron.

@@

\paragraph{Fiducial cuts}
~\\
~\\
~~~~Some regions of CLAS have to be excluded from the analysis to ensure an accurate detection of the different particles. For instance, an electron that hits the edge of the EC will have only part of its electromagnetic shower contained within the detector. Also, the structure of the torus magnet divides CLAS into six separate sectors, which makes edge effects non-negligible. For this reason, the following set of fiducial cuts is applied:

@@

\item CC fiducial cut: in reference \cite{CCref}, G. Adams {\it et al.} have studied the efficiency of the CCs. They found that within the fiducial regions, which are defined by the edges of the CCs mirrors, the detection efficiency is stable and is around 98$\%$. Outside the fiducial regions, the efficiency shows strong variations. In reference \cite{Osipenko}, M. Osipenko {\it et al.} have developed a coordinate system to represent the CC hits with respect to the center of each sector in CLAS. In this frame, the CCs mirror edges can be defined as:\\
\begin{multline}  
\phi_{e}  = -63.32792 +11.05609\cdot \theta_{e} -0.6344957\cdot \theta^{2}_{e} +1.873895*10^{-2}\cdot \theta^{3}_{e} \\
- 2.762131*10^{-2}\cdot \theta^{4}_{e}+1.604035*10^{-2}\cdot \theta^{5}_{e}. ~~~~~~~~~~~~~~~~~~~~~~~~~
\end{multline}
Based on these works, we reject all the hits located outside the mirror edges. Figure \ref{fig:CC_cut} shows an illustration of this cut.

@@

\paragraph{EC energy cuts}
~\\
~\\
~~~~The Minimum Ionizing particles (MIPs), such as pions, deposit constant 
amounts of energy per distance while traversing the EC.  In contrast, the 
showering particles, such as electrons and photons, deposit energies 
proportional to their momenta. We use two energy cuts to clean the electrons 
from the main contamination, i.e.  $\pi^{-}$s.
\begin{itemize}
\item Minimum deposited energy:
   the inner and the outer parts of the EC have thicknesses 15 cm and 24 cm, 
   respectively. The simulations show that pions deposit a constant energy 
   amount of 2 MeV/cm, independently of their momenta. Figure 
   \ref{fig:Einner_cut} shows the deposited energy in the outer part of the EC 
   (EC$_{out}$) as a function of the energy deposited in the inner part 
   (EC$_{in}$), after the fiducial cuts. On the x-axis, one can see a clear 
   region, around 30 MeV, that comes mainly from the negative pions, which 
   deposit 2 MeV/cm along the 15 cm thickness of the inner EC. We use a cut of 
   60 MeV on EC$_{in}$ to reject these particles.

@@

\paragraph{CC cut}
~\\
~\\
~~~~The Cherenkov counters have been designed to separate electrons from pions below 2.5 GeV/c momentum. In this region, the pions are not supposed to produce photoelectrons. However, low momentum $\delta$-electrons can be produced from the diffusion of the pions in the Cherenkov gas. These $\delta$-electrons produce a small number of photoelectrons. Figure \ref{fig:Nphe_cut} shows the distributions of the number of photoelectrons ($nphe$) produced by the negative particles for three different stages of the electron selection. \\

@@

\subsection{Proton identification}
~~~~Similarly to the electrons, the protons are affected by the geometry and the efficy of the different sub-detectors of CLAS. The following conditions are applied to select the good protons.
\begin{itemize}
\item Coincidence with one and only one good electron.

@@

\item Vertex matching:  the last cut is the correspondence between the longitudinal vertices of the detected electron and proton. Figure \ref{fig:proton_delta_z_e_prot} shows the difference $\Delta z = z_{e} - z_{p}$ and the chosen cuts (red dashed lines).
\end{itemize}
~~~~Finally, figure \ref{fig:prot_Theta_Phi} shows the azimuthal angle as a function of the polar angle distribution for the identified protons after all the selection cuts. One notices that the population of the protons is different from one sector to another, which comes from the dead regions in some sectors.
\vspace{-0.1in}

@@

\subsection{Photon identification}
~~~~CLAS is equipped with two calorimeters that can detect photons: the IC, covering polar angles from 4$^\circ$ to 14$^\circ$, and the EC, covering polar angles from 8$^\circ$ to 45$^\circ$. Like for the other particles, in addition to the coincidence with one good electron, we require a set of criteria to ensure the quality of the detected photons. Due to the efficiency constraints in both calorimeters, we restrict the energy of the selected photons to be greater than 300 MeV. Further requirements are applied depending on each detector.

@@

\subsection{Helium-4 identification}
~~~~Details on the RTPC's structure and the definition of the different 
variables were given in chapter 2, section 2. In the following, we show the 
cuts to select the RTPC good tracks with a 6-GeV beam energy. The distributions 
will be shown for the two independent modules of the RTPC separately, labelled 
as left and right sides. The cuts are:

@@

\subsection{$\pi^{0}$ identification}
~~~~In this analysis, we identify the $\pi^{0}$s with the goal of DVCS background subtraction as will be addressed in section \ref{Background_subtraction}. The $\pi^{0}$ identification is based on its two real photons decay mode ($\pi^{0} \rightarrow \gamma \gamma$), with a branching ratio around 98.8$\%$. In our experimental configuration, the reconstructed neutral pions can be categorized into three topologies: ICIC (the two photons are detected in the IC), ICEC (one photon in the IC and the second in the EC), and ECEC (the two photons are detected in the EC). The nominal mass of the $\pi^{0}$ is 0.135 GeV/c$^{2}$. The reconstructed invariant mass of the detected pair of photons can be seen in figure \ref{fig:pi0_selection}, for the three topologies. One can see clear peaks corresponding to the neutral pions in the three distributions.
\begin{figure}[tbp]
\hspace{-0.2in}
\includegraphics[scale=0.34]{fig_analysis/m_pi0_icic.png}
\hspace{-0.2in}
\includegraphics[scale=0.34]{fig_analysis/m_pi0_ecic.png}
\hspace*{-0.1in}
\includegraphics[scale=0.34]{fig_analysis/m_pi0_ecec.png}
\hspace{-0.2in}
\includegraphics[scale=0.38]{fig_analysis/sim_photon_phi_theta_2.png}
\caption{The reconstructed invariant mass of the photon pairs in the three topologies: ICIC, ECIC and ECEC. On the bottom-right: $\phi$ vs. $\theta$ distribution for the simulated DVCS photons, shown that most of the DVCS photons are located in the IC.} 
\label{fig:pi0_selection}
\end{figure}\\

@@

\section{Simulation}
~~~~The Monte-Carlo simulation is used in this analysis for two goals: understanding the behaviour of the particles of interest in the detectors, and computing the acceptance for the DVCS background subtraction. The key stages of the simulation are summarized in figure \ref{fig:simulation_steps}: Monte-Carlo generated events pass in a GEANT3 simulation (GSIM) for the CLAS detector, then are processed with a package called GPP to add resolution and efficiency effects, and finally are reconstructed by a package named RECSIS, from which we obtain the physical quantities of each particle, like it is done for real data. In the following, each simulation stage will be briefly presented.
\begin{figure}[tbp]
\centering
\includegraphics[scale=0.25]{fig_simulation/simulation_steps.png}
\caption{ The scheme shows the main simulation steps.} 
\label{fig:simulation_steps}
\end{figure}

\subsection{Event generator} \label{Event_generator}
~~~~Events are generated in the measured ranges of $Q^{2}$, $x_{B}$, $t$, and 
$\phi$ following a parametrization of the cross section which roughly 
reproduces the DVCS and exclusive $\pi^{0}$ electroproduction data 
\cite{FX_thesis}:
\begin{equation}
\frac{d^{4}\sigma}{dQ^{2}dx_{B}dtd\phi} \propto \left(\frac{Q^{2}_{0}}{Q^{2}}\right)^{\alpha} *  \frac{1}{1 + (\frac{x_{B} - x_{c}}{c})^{2}} * \frac{1}{(1+bt)^{\beta}} * (1-d(1-cos(\phi)).
\label{equ:event_generator}
\end{equation}
This parametrization is the product of four factors which reproduce the DVCS and $\pi^{0}$,  characteristics as follows:
\begin{itemize}
\item the $Q^{2}$-dependent term accounts for the depth of the interaction: $Q^{2}_{0}$ is the minimum allowed value and the $\alpha$ is a parameter which controls the shape of the distribution.   
\item the $x_{B}$ term accounts for the dependence of the cross section on the parton distribution functions, with $x_{c}$ the mean value of the Bjorken variable $x_{B}$.
\item the $t$ term accounts for the t-dependence of the elastic form factors of the helium and of the proton, via the parameters $b$ and $\beta$.
\item the $\phi$ term accounts for the cross section dependence on this angle, via the parameter $d$. The DVCS and the BH have a different behaviour than the $\pi^{0}$ exclusive events. 
\end{itemize}

~~~~Table \ref{Table:event_generator_values} shows the values of the parameters used for the cross section parametrization of the four channels of interest: $e^{4}He\gamma$, $e^{4}He\pi^{0}$, $ep\gamma$, and $ep\pi^{0}$.

@@

~~~~Regarding the protons in $ep\gamma$ and $ep\pi^{0}$ channels, we apply the 
Fermi motion on the initial protons based on the parametrization of C. Ciofi 
degli Atii and S. Simula \cite{fermi_motion}. Figure \ref{fig:fermi_mom} shows 
the Fermi momentum distributions of the nucleons inside $^{4}He$, where we 
apply a cut on the high-momentum tail at 0.3 GeV/c, higher momentum should not 
contribute to the final DVCS sample because of our exclusive cuts presented in 
the following chapter.

@@

\subsection{GSIM}
~~~~GSIM is a GEANT3-based simulation of the CLAS detector, developed by the CLAS collaboration \cite{GSIM}. This simulation takes into account all the experimental environment, such as the target's position, materials, and geometry, and also the various sub-detectors' materials and geometries, to reproduce the behaviour of particles traversing the detector. {\it Nota Bene} the reconstruction in the RTPC is not implemented in GSIM, however its material is present.

\subsection{GPP}
~~~~For the purpose of making the simulation more realistic, the output of GSIM is fed to the GSIM Post-Processing (GPP) package. This package applies resolution effects on the different measured quantities, and also reads efficiency maps to match the simulation to the real experiment. The resolutions are categorized into three groups, time, position, and energy resolution. In the following, we present the techniques used in this work to extract the smearing factors for the simulation.

\paragraph{SC time smearing}
~\\
~\\
~~~~The GPP takes a single factor to smear the time in a Gaussian form. Figure \ref{fig:SC_smearing_factors} shows an illustration of the time smearing effects, in which $\Delta \beta$ for the protons \\($\Delta \beta~=~\beta_{measured}~-~\beta_{calculated}$~) is compared for data and simulation. The chosen SC smearing factor is 2.1.

\paragraph{DCs position smearing}
~\\
~\\
~~~~The GPP uses three factors to smear the positions of the hits in the DCs, in a Gaussian form. These smearing factors are extracted from comparing simulation to data in terms of the Time-Based Tracking (TBT) residual distributions, i.e. the deviation of the hits in the DCs from the fitted track. Figure \ref{fig:DC_smearing_factors} shows the experimental, initial simulated, and smeared simulated residual distributions for the collected good electrons and protons. The extracted smearing factors are: ($a$, $b$, $c$) = (1.1, 0.85, 1.1), where the factor $a$ stands for smearing of DC1, $b$ for DC2, and $c$ for DC3. One can see the smearing effects by looking to the values of $\sigma$ from the fit in each Super-Layer (SL). We conclude that these factors match the simulation to the data.
\begin{figure}[tbp]
\hspace{-0.7in}
\includegraphics[scale=0.3]{fig_simulation/Exp_proton_delta_beta.png}
\hspace{-0.1in}
\includegraphics[scale=0.32]{fig_simulation/Sim_proton_delta_beta.png}
\centering
\includegraphics[scale=0.32]{fig_simulation/Sim_proton_delta_beta_smeared.png}
\caption{Illustration of the SC time smearing via the $\Delta \beta$ distributions for the protons. On the top left: $\Delta \beta$ distribution of the experimental data. On the top right: the simulated data without smearing. On the bottom: the simulated data after SC time smearing with a factor equal to 2.1.} 
\label{fig:SC_smearing_factors}
\end{figure}

@@

\paragraph{IC energy smearing}
~\\
~\\
~~~~An energy smearing for the IC photons is also needed. The GPP software uses three factors to perform this smearing: $ic\_a$, $ic\_b$, and $ic\_c$, where $ic\_a$ accounts for the smearing of the width of the noise around zero using a Gaussian, while $ic\_b$ and $ic\_c$ are the smearing factors for the ADC values, using Gaussians as well. Our parameters are: 0.008, 0.036, and 0.024 respectively. Figure \ref{fig:IC_smearing_factors} shows the effect of these smearing factors on the simulated invariant mass of the $\pi^{0}$. The associated experimental distribution can be seen on the top left plot of figure \ref{fig:pi0_selection}. One notices that the smeared width of the simulated distribution matches the measured experimental one.\\

@@

~~~~RECSIS is the reconstruction package of CLAS which is used for both data 
and simulation. The detector responses, in terms of ADCs and TDCs, are 
converted by RECSIS into physical meaningful physical quantities, such as 
momentum, using lookup tables. In this process, the thresholds on the different 
detectors (DCs, CCs, ECs, SCs, IC, and RTPC) are applied to filter the signals 
before the reconstruction procedures. After this reconstruction, the outputs of 
RECSIS for the experimental and the simulated events can be compared directly, 
as will be presented in the next section.

\subsection{RTPC fastmc}
~~~~All the sub-detectors of CLAS are implemented in the GSIM GEANT3 simulation, 
however the RTPC is only there as dead material. As it is a more recent detector
it has been fully simulated using a stand alone GEANT4 simulation. As a 
first step, we replace this RTPC's simulation by a fast Monte-Carlo (fastmc) 
package at the level of the event generator to make the simulated $^{4}He$ more 
realistic and matching the data. This fastmc smears the $^{4}He$ kinematics and 
applies the RTPC's acceptance. Regarding the smearing, the momentum, polar 
angle, azimuthal angle and z-vertex of the $^{4}He$ are smeared with Gaussians 
using the observed tracking resolutions of the RTPC (see chapter 2, section 3).  
For the acceptance, the fastmc:
\begin{itemize}
\item ensures that the $^{4}He$ track intersects the cathode and the gems of the RTPC.
\item removes the tracks which pass in the dead area between the two modules of the RTPC.
\item removes the track if it goes to the upstream end of the target's holder.
\item applies the RTPC's thresholds on the momentum and the polar angle.
\end{itemize}

@@

\section{Kinematic Corrections}
~~~~~The simulation enables us to extract kinematic corrections for the different particles, by comparing the reconstructed physical quantities from the simulation to the initially generated ones. In general, the reconstructed azimuthal angles of all the particles are consistent with the generated ones, while the reconstructed polar angles and momenta show some deviations. Therefore, corrections are required. In the following, the electron and the proton corrections are applied on both data and simulation, while the simulated photons are corrected differently from the experimental ones, as will be explained and justified.

\subsection{Electrons}
~~~~The reconstructed kinematics of the simulated electrons are roughly consistent with the generated quantities, as can be seen in figure \ref{fig:electron_corrections}. Nevertheless, we extracted and applied corrections on the electrons to achieve a higher precision. The corrected polar angles and momenta take the form:
\begin{equation}
\theta_{corr.} = \frac{\theta_{rec.}}{f(\theta_{rec.})} ~~~~~~~ p_{corr.} = \frac{p_{rec.}}{f(p_{rec.})}
\label{equation:correcting_function}
\end{equation}  
where $f(\theta_{Rec.})$ and $f(p_{Rec.})$ are the functions shown, respectively, on the left and right plots of figure \ref{fig:electron_corrections}.
\begin{figure}[tbp]
\centering
\includegraphics[scale=0.41]{fig_simulation/before_electron_Rec_gen.png}
\caption{Electron corrections. On the left: The ratio of the reconstructed to the generated electron polar angles is plotted as a function of the reconstructed polar angle. The mean of the distribution is parametrized by the black curve. On the right: the momentum ratio ($Rec/Gen$) is plotted as a function of the reconstructed momentum and parametrized by the black curve .} 
\label{fig:electron_corrections}
\end{figure}

\subsection{Protons}
~~~~The same procedures for the extraction of the kinematic corrections of the electrons have been carried out for the protons. The results can be seen in figure \ref{fig:proton_corrections}. The corrected polar angles and momenta are obtained using equations \ref{equation:correcting_function} and the functions shown on each plot of figure \ref{fig:proton_corrections}. 
\begin{figure}[tp]
\centering
\includegraphics[scale=0.41]{fig_simulation/before_proton_Rec_gen.png}
\caption{Proton corrections. On the left: the reconstructed-generated polar angle ratio is plotted as a function of the reconstructed angle and fitted with the black curve. On the right: the momentum ratio is plotted as a function of the reconstructed momentum.} 
\label{fig:proton_corrections}
\end{figure}

>>>>

\chapter{Particles reconstruction and simulation}
The final state of a coherent (incoherent) DVCS event consists of three 
particles: an electron, a $^4$He (a proton), and a real photon. To identify the 
DVCS events, we first identify, individually, the different particles of 
interest. Then, events with three detected final-state particles will be 
further filtered by imposing the energy-momentum conservation laws, as will be 
presented in the following chapter.

Even after imposing the conservation laws, the DVCS sample will not have 
100$\%$ truly DVCS events. In our kinematical region, the main contamination to 
our DVCS channels comes from the electroproduction of neutral pions. For 
instance, in the coherent $\pi^0$-electroproduction, when one of the two-photon 
decay of the $\pi^{0}$ passes the DVCS requirements, it will be counted as a 
DVCS event. Thus, these events have to be subtracted before looking to any DVCS 
observable. For this purpose, we perform a technique in which we combine the 
measured exclusive $\pi^0$-electroproduction data sample with a Monte-Carlo 
simulation to evaluate the background in the selected DVCS sample. For this 
technique, we need to identify the experimental $\pi^0$s that come in the 
coherent and the incoherent $\pi^0$-electroproduction channels.

@@

\subsection{Electron identification} \label{Electron_identification}
The electron detection triggers the data acquisition system to record data 
from all the sub-detectors of CLAS. In this analysis, a particle which passes 
the following set of criteria is assumed to be a good electron.

@@

\paragraph{Fiducial cuts} ~\\
Some regions of CLAS have to be excluded from the analysis to ensure an accurate detection of the different particles. For instance, an electron that hits the edge of the EC will have only part of its electromagnetic shower contained within the detector. Also, the structure of the torus magnet divides CLAS into six separate sectors, which makes edge effects non-negligible. For this reason, the following set of fiducial cuts is applied:

@@

\item CC fiducial cut: in reference \cite{CCref}, G. Adams {\it et al.} have studied the efficiency of the CCs. They found that within the fiducial regions, which are defined by the edges of the CCs mirrors, the detection efficiency is stable and is around 98$\%$. Outside the fiducial regions, the efficiency shows strong variations. In reference \cite{Osipenko}, M. Osipenko {\it et al.} have developed a coordinate system to represent the CC hits with respect to the center of each sector in CLAS. In this frame, the CCs mirror edges can be defined as:
\begin{multline}  
\phi_{e}  = -63.32792 +11.05609\cdot \theta_{e} -0.6344957\cdot \theta^{2}_{e} +1.873895*10^{-2}\cdot \theta^{3}_{e} \\
- 2.762131*10^{-2}\cdot \theta^{4}_{e}+1.604035*10^{-2}\cdot \theta^{5}_{e}. ~~~~~~~~~~~~~~~~~~~~~~~~~
\end{multline}
Based on these works, we reject all the hits located outside the mirror edges. Figure \ref{fig:CC_cut} shows an illustration of this cut.

@@

\paragraph{EC energy cuts} ~\\
The Minimum Ionizing particles (MIPs), such as pions, deposit constant 
amounts of energy per distance while traversing the EC.  In contrast, the 
showering particles, such as electrons and photons, deposit energies 
proportional to their momenta. We use two energy cuts to clean the electrons 
from the main contamination, i.e.  $\pi^{-}$s.
\begin{itemize}
\item Minimum deposited energy:
   the inner and the outer parts of the EC have thicknesses 15 cm and 24 cm, 
   respectively. The simulations show that pions deposit a constant energy 
   amount of 2 MeV/cm, independently of their momenta. Figure 
   \ref{fig:Einner_cut} shows the deposited energy in the outer part of the EC 
   (EC$_{out}$) as a function of the energy deposited in the inner part 
   (EC$_{in}$), after the fiducial cuts. On the x-axis, one can see a clear 
   region, around 30 MeV, that comes mainly from the negative pions, which 
   deposit 2 MeV/cm along the 15 cm thickness of the inner EC. We use a cut of 
   60 MeV on EC$_{in}$ to reject these particles.

@@

\paragraph{CC cut} ~\\
The Cherenkov counters have been designed to separate electrons from pions below 2.5 GeV/c momentum. In this region, the pions are not supposed to produce photoelectrons. However, low momentum $\delta$-electrons can be produced from the diffusion of the pions in the Cherenkov gas. These $\delta$-electrons produce a small number of photoelectrons. Figure \ref{fig:Nphe_cut} shows the distributions of the number of photoelectrons ($nphe$) produced by the negative particles for three different stages of the electron selection.

@@

\subsection{Proton identification}
Similarly to the electrons, the protons are affected by the geometry and the efficy of the different sub-detectors of CLAS. The following conditions are applied to select the good protons.
\begin{itemize}
\item Coincidence with one and only one good electron.

@@

\item Vertex matching:  the last cut is the correspondence between the longitudinal vertices of the detected electron and proton. Figure \ref{fig:proton_delta_z_e_prot} shows the difference $\Delta z = z_{e} - z_{p}$ and the chosen cuts (red dashed lines).
\end{itemize}
Finally, figure \ref{fig:prot_Theta_Phi} shows the azimuthal angle as a function of the polar angle distribution for the identified protons after all the selection cuts. One notices that the population of the protons is different from one sector to another, which comes from the dead regions in some sectors.
\vspace{-0.1in}

@@

\subsection{Photon identification}
CLAS is equipped with two calorimeters that can detect photons: the IC, covering polar angles from 4$^\circ$ to 14$^\circ$, and the EC, covering polar angles from 8$^\circ$ to 45$^\circ$. Like for the other particles, in addition to the coincidence with one good electron, we require a set of criteria to ensure the quality of the detected photons. Due to the efficiency constraints in both calorimeters, we restrict the energy of the selected photons to be greater than 300 MeV. Further requirements are applied depending on each detector.

@@

\subsection{Helium-4 identification}
Details on the RTPC's structure and the definition of the different 
variables were given in chapter 2, section 2. In the following, we show the 
cuts to select the RTPC good tracks with a 6-GeV beam energy. The distributions 
will be shown for the two independent modules of the RTPC separately, labelled 
as left and right sides. The cuts are:

@@

\subsection{$\pi^{0}$ identification}
In this analysis, we identify the $\pi^{0}$s with the goal of DVCS background subtraction as will be addressed in section \ref{Background_subtraction}. The $\pi^{0}$ identification is based on its two real photons decay mode ($\pi^{0} \rightarrow \gamma \gamma$), with a branching ratio around 98.8$\%$. In our experimental configuration, the reconstructed neutral pions can be categorized into three topologies: ICIC (the two photons are detected in the IC), ICEC (one photon in the IC and the second in the EC), and ECEC (the two photons are detected in the EC). The nominal mass of the $\pi^{0}$ is 0.135 GeV/c$^{2}$. The reconstructed invariant mass of the detected pair of photons can be seen in figure \ref{fig:pi0_selection}, for the three topologies. One can see clear peaks corresponding to the neutral pions in the three distributions.
\begin{figure}[tbp]
\hspace{-0.2in}
\includegraphics[scale=0.34]{fig_analysis/m_pi0_icic.png}
\hspace{-0.2in}
\includegraphics[scale=0.34]{fig_analysis/m_pi0_ecic.png}
\hspace*{-0.1in}
\includegraphics[scale=0.34]{fig_analysis/m_pi0_ecec.png}
\hspace{-0.2in}
\includegraphics[scale=0.38]{fig_analysis/sim_photon_phi_theta_2.png}
\caption{The reconstructed invariant mass of the photon pairs in the three topologies: ICIC, ECIC and ECEC. On the bottom-right: $\phi$ vs. $\theta$ distribution for the simulated DVCS photons, shown that most of the DVCS photons are located in the IC.} 
\label{fig:pi0_selection}
\end{figure}

@@

\section{Simulation}
The Monte-Carlo simulation is used in this analysis for two goals: understanding the behaviour of the particles of interest in the detectors, and computing the acceptance for the DVCS background subtraction. The key stages of the simulation are summarized in figure \ref{fig:simulation_steps}: Monte-Carlo generated events pass in a GEANT3 simulation (GSIM) for the CLAS detector, then are processed with a package called GPP to add resolution and efficiency effects, and finally are reconstructed by a package named RECSIS, from which we obtain the physical quantities of each particle, like it is done for real data. In the following, each simulation stage will be briefly presented.
\begin{figure}[tbp]
\centering
\includegraphics[scale=0.25]{fig_simulation/simulation_steps.png}
\caption{ The scheme shows the main simulation steps.} 
\label{fig:simulation_steps}
\end{figure}

\subsection{Event generator} \label{Event_generator}
Events are generated in the measured ranges of $Q^{2}$, $x_{B}$, $t$, and 
$\phi$ following a parametrization of the cross section which roughly 
reproduces the DVCS and exclusive $\pi^{0}$ electroproduction data 
\cite{FX_thesis}:
\begin{equation}
\frac{d^{4}\sigma}{dQ^{2}dx_{B}dtd\phi} \propto \left(\frac{Q^{2}_{0}}{Q^{2}}\right)^{\alpha} *  \frac{1}{1 + (\frac{x_{B} - x_{c}}{c})^{2}} * \frac{1}{(1+bt)^{\beta}} * (1-d(1-cos(\phi)).
\label{equ:event_generator}
\end{equation}
This parametrization is the product of four factors which reproduce the DVCS and $\pi^{0}$,  characteristics as follows:
\begin{itemize}
\item the $Q^{2}$-dependent term accounts for the depth of the interaction: $Q^{2}_{0}$ is the minimum allowed value and the $\alpha$ is a parameter which controls the shape of the distribution.   
\item the $x_{B}$ term accounts for the dependence of the cross section on the parton distribution functions, with $x_{c}$ the mean value of the Bjorken variable $x_{B}$.
\item the $t$ term accounts for the t-dependence of the elastic form factors of the helium and of the proton, via the parameters $b$ and $\beta$.
\item the $\phi$ term accounts for the cross section dependence on this angle, via the parameter $d$. The DVCS and the BH have a different behaviour than the $\pi^{0}$ exclusive events. 
\end{itemize}

Table \ref{Table:event_generator_values} shows the values of the parameters used for the cross section parametrization of the four channels of interest: $e^{4}He\gamma$, $e^{4}He\pi^{0}$, $ep\gamma$, and $ep\pi^{0}$.

@@

Regarding the protons in $ep\gamma$ and $ep\pi^{0}$ channels, we apply the 
Fermi motion on the initial protons based on the parametrization of C. Ciofi 
degli Atii and S. Simula \cite{fermi_motion}. Figure \ref{fig:fermi_mom} shows 
the Fermi momentum distributions of the nucleons inside $^{4}He$, where we 
apply a cut on the high-momentum tail at 0.3 GeV/c, higher momentum should not 
contribute to the final DVCS sample because of our exclusive cuts presented in 
the following chapter.

@@

\subsection{GSIM}
GSIM is a GEANT3-based simulation of the CLAS detector, developed by the CLAS collaboration \cite{GSIM}. This simulation takes into account all the experimental environment, such as the target's position, materials, and geometry, and also the various sub-detectors' materials and geometries, to reproduce the behaviour of particles traversing the detector. {\it Nota Bene} the reconstruction in the RTPC is not implemented in GSIM, however its material is present.

\subsection{GPP}
For the purpose of making the simulation more realistic, the output of GSIM is fed to the GSIM Post-Processing (GPP) package. This package applies resolution effects on the different measured quantities, and also reads efficiency maps to match the simulation to the real experiment. The resolutions are categorized into three groups, time, position, and energy resolution. In the following, we present the techniques used in this work to extract the smearing factors for the simulation.

\paragraph{SC time smearing} ~\\
The GPP takes a single factor to smear the time in a Gaussian form. Figure \ref{fig:SC_smearing_factors} shows an illustration of the time smearing effects, in which $\Delta \beta$ for the protons ($\Delta\beta~=~\beta_{measured}-\beta_{calculated}$) is compared for data and simulation. The chosen SC smearing factor is 2.1.

\paragraph{DCs position smearing} ~\\
The GPP uses three factors to smear the positions of the hits in the DCs, in a Gaussian form. These smearing factors are extracted from comparing simulation to data in terms of the Time-Based Tracking (TBT) residual distributions, i.e. the deviation of the hits in the DCs from the fitted track. Figure \ref{fig:DC_smearing_factors} shows the experimental, initial simulated, and smeared simulated residual distributions for the collected good electrons and protons. The extracted smearing factors are: ($a$, $b$, $c$) = (1.1, 0.85, 1.1), where the factor $a$ stands for smearing of DC1, $b$ for DC2, and $c$ for DC3. One can see the smearing effects by looking to the values of $\sigma$ from the fit in each Super-Layer (SL). We conclude that these factors match the simulation to the data.
\begin{figure}[tbp]
\hspace{-0.7in}
\includegraphics[scale=0.3]{fig_simulation/Exp_proton_delta_beta.png}
\hspace{-0.1in}
\includegraphics[scale=0.32]{fig_simulation/Sim_proton_delta_beta.png}
\centering
\includegraphics[scale=0.32]{fig_simulation/Sim_proton_delta_beta_smeared.png}
\caption{Illustration of the SC time smearing via the $\Delta \beta$ distributions for the protons. On the top left: $\Delta \beta$ distribution of the experimental data. On the top right: the simulated data without smearing. On the bottom: the simulated data after SC time smearing with a factor equal to 2.1.} 
\label{fig:SC_smearing_factors}
\end{figure}

@@

\paragraph{IC energy smearing} ~\\
An energy smearing for the IC photons is also needed. The GPP software uses three factors to perform this smearing: $ic\_a$, $ic\_b$, and $ic\_c$, where $ic\_a$ accounts for the smearing of the width of the noise around zero using a Gaussian, while $ic\_b$ and $ic\_c$ are the smearing factors for the ADC values, using Gaussians as well. Our parameters are: 0.008, 0.036, and 0.024 respectively. Figure \ref{fig:IC_smearing_factors} shows the effect of these smearing factors on the simulated invariant mass of the $\pi^{0}$. The associated experimental distribution can be seen on the top left plot of figure \ref{fig:pi0_selection}. One notices that the smeared width of the simulated distribution matches the measured experimental one.

@@

RECSIS is the reconstruction package of CLAS which is used for both data 
and simulation. The detector responses, in terms of ADCs and TDCs, are 
converted by RECSIS into physical meaningful physical quantities, such as 
momentum, using lookup tables. In this process, the thresholds on the different 
detectors (DCs, CCs, ECs, SCs, IC, and RTPC) are applied to filter the signals 
before the reconstruction procedures. After this reconstruction, the outputs of 
RECSIS for the experimental and the simulated events can be compared directly, 
as will be presented in the next section.

\subsection{RTPC fastmc}
All the sub-detectors of CLAS are implemented in the GSIM GEANT3 simulation, 
however the RTPC is only there as dead material. As it is a more recent detector
it has been fully simulated using a stand alone GEANT4 simulation. As a 
first step, we replace this RTPC's simulation by a fast Monte-Carlo (fastmc) 
package at the level of the event generator to make the simulated $^{4}He$ more 
realistic and matching the data. This fastmc smears the $^{4}He$ kinematics and 
applies the RTPC's acceptance. Regarding the smearing, the momentum, polar 
angle, azimuthal angle and z-vertex of the $^{4}He$ are smeared with Gaussians 
using the observed tracking resolutions of the RTPC (see chapter 2, section 3).  
For the acceptance, the fastmc:
\begin{itemize}
\item ensures that the $^{4}He$ track intersects the cathode and the gems of the RTPC.
\item removes the tracks which pass in the dead area between the two modules of the RTPC.
\item removes the track if it goes to the upstream end of the target's holder.
\item applies the RTPC's thresholds on the momentum and the polar angle.
\end{itemize}

@@

\section{Kinematic Corrections}
The simulation enables us to extract kinematic corrections for the different particles, by comparing the reconstructed physical quantities from the simulation to the initially generated ones. In general, the reconstructed azimuthal angles of all the particles are consistent with the generated ones, while the reconstructed polar angles and momenta show some deviations. Therefore, corrections are required. In the following, the electron and the proton corrections are applied on both data and simulation, while the simulated photons are corrected differently from the experimental ones, as will be explained and justified.

\subsection{Electrons}
The reconstructed kinematics of the simulated electrons are roughly consistent with the generated quantities, as can be seen in figure \ref{fig:electron_corrections}. Nevertheless, we extracted and applied corrections on the electrons to achieve a higher precision. The corrected polar angles and momenta take the form:
\begin{equation}
\theta_{corr.} = \frac{\theta_{rec.}}{f(\theta_{rec.})} ~~~~~~~ p_{corr.} = \frac{p_{rec.}}{f(p_{rec.})}
\label{equation:correcting_function}
\end{equation}  
where $f(\theta_{Rec.})$ and $f(p_{Rec.})$ are the functions shown, respectively, on the left and right plots of figure \ref{fig:electron_corrections}.
\begin{figure}[tbp]
\centering
\includegraphics[scale=0.41]{fig_simulation/before_electron_Rec_gen.png}
\caption{Electron corrections. On the left: The ratio of the reconstructed to the generated electron polar angles is plotted as a function of the reconstructed polar angle. The mean of the distribution is parametrized by the black curve. On the right: the momentum ratio ($Rec/Gen$) is plotted as a function of the reconstructed momentum and parametrized by the black curve .} 
\label{fig:electron_corrections}
\end{figure}

\subsection{Protons}
The same procedures for the extraction of the kinematic corrections of the electrons have been carried out for the protons. The results can be seen in figure \ref{fig:proton_corrections}. The corrected polar angles and momenta are obtained using equations \ref{equation:correcting_function} and the functions shown on each plot of figure \ref{fig:proton_corrections}. 
\begin{figure}[tp]
\centering
\includegraphics[scale=0.41]{fig_simulation/before_proton_Rec_gen.png}
\caption{Proton corrections. On the left: the reconstructed-generated polar angle ratio is plotted as a function of the reconstructed angle and fitted with the black curve. On the right: the momentum ratio is plotted as a function of the reconstructed momentum.} 
\label{fig:proton_corrections}
\end{figure}
